---
title: "Bootstrap Vignette"
output: rmarkdown::pdf_document
vignette: >
  %\VignetteIndexEntry{Example Vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
bibliography: bibliography.bib
nocite: "@*"
---

```{r include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE
)
knitr::opts_chunk$set(fig.width=7, fig.height=6)
```

```{r}
library("STA380project")
```

This vignette demonstrates three functions for sleep duration statistics with the bootstrap resampling method:

-   compute_sleep_duration(): Compute sleep duration as the difference between wake time and bedtime.

-   bootstrap_mean(): Generate bootstrap estimates of the sample mean using resampling with replacement.

-   bootstrap_ci() Compute a percentile bootstrap confidence interval.

# Compute sleep duration

In this example, bedtimes and wake times are recorded on a 24-hour clock. If wake time is smaller than bedtime, the function assumes sleep crosses midnight and adjusts accordingly.

```{r}
set.seed(1)
n <- 10^4
uptime  <- c( 7.0, 6.0, 9.0, 7.5, 6.0)
bedtime <- c(22.0, 22.5, 1.0, 0.0, 0.5)
sleep_hours <- compute_sleep_duration(bedtime, uptime)
sleep_hours
summary(sleep_hours)
```

# Bootstrap mean of sleep duration

we apply the bootstrap to estimate the mean sleep duration of the sampling distribution.

```{r}
set.seed(1)
B <- 10^4
boot_means_sleep <- bootstrap_mean(sleep_hours, B = B)
head(boot_means_sleep)
length(boot_means_sleep)
```

## Histogram

Visualizing the bootstrap of mean sleep duration

```{r}
hist(boot_means_sleep,
     main = "Bootstrap distribution of mean sleep duration",
     xlab = "Bootstrap means(hours)")
```

## Bootstrap summary statistics (Mean)

```{r}
original_mean <- mean(sleep_hours)

boot_mean <- mean(boot_means_sleep)
boot_se <- sd(boot_means_sleep)
boot_bias <- boot_mean - original_mean
boot_median <- median(boot_means_sleep)
boot_ci <- quantile(boot_means_sleep, c(0.025, 0.975))

mean_summary <- data.frame(
  Metric = c("Original estimate",
             "Bootstrap mean",
             "Bootstrap SE",
             "Bias",
             "Bootstrap median",
             "95% CI (2.5%)",
             "95% CI (97.5%)"),
  Value = c(original_mean,
            boot_mean,
            boot_se,
            boot_bias,
            boot_median,
            boot_ci[1],
            boot_ci[2])
)

mean_summary
```

# Bootstrap Median Estimator

We apply the bootstrap to estimate the median sleep duration of the sampling distribution. 

```{r}
set.seed(1)
B <- 10^4
boot_median_sleep <- bootstrap_median(sleep_hours, B = B)
head(boot_median_sleep)
length(boot_median_sleep)
```

## histogram

Visualizing the bootstrap of mean sleep duration.

```{r}
hist(boot_median_sleep,
     main = "Bootstrap distribution of median sleep duration",
     xlab = "Bootstrap median(hours)")
```

## Bootstrap summary statistics (Median)

```{r}
original_median <- median(sleep_hours)

boot_mean_med <- mean(boot_median_sleep)
boot_se_med <- sd(boot_median_sleep)
boot_bias_med <- boot_mean_med - original_median
boot_median_med <- median(boot_median_sleep)
boot_ci_med <- quantile(boot_median_sleep, c(0.025, 0.975))

median_summary <- data.frame(
  Metric = c("Original estimate",
             "Bootstrap mean",
             "Bootstrap SE",
             "Bias",
             "Bootstrap median",
             "95% CI (2.5%)",
             "95% CI (97.5%)"),
  Value = c(original_median,
            boot_mean_med,
            boot_se_med,
            boot_bias_med,
            boot_median_med,
            boot_ci_med[1],
            boot_ci_med[2])
)

median_summary
```

# Percentile bootstrap confidence interval

Compute the lower and upper bound of the 95% percentile bootstrap confidence interval.

```{r}
cl95_sleep <- bootstrap_ci(boot_means_sleep, conf_level = 0.95)
cl95_sleep
```

# Assumption checks

In this bootstrap, we do not need to check normality, but it needs the following assumptions:

-   Independence: Observations of sleep duration should be independent.
-   Representative: Observations should come from the target population.
-   Sample size: Sample size n should be big enough to make bootstrap_ci() stable.
-   Outliers: Outliers will effect the bootstrap_mean() and bootstrap_ci().

# Checking sample size

```{r}
sleep_sample <- sleep_hours[!is.na(sleep_hours)]
n <- length(sleep_sample)
if (n < 10) {
  warning("Sample size n is too small!")
}
```

# Checking outliers

```{r}
sleep_sample <- sleep_hours[!is.na(sleep_hours)]
Q1 <- quantile(sleep_sample, 0.25)
Q3 <- quantile(sleep_sample, 0.75)
IQR <- IQR(sleep_sample)
lower_tail <- Q1 - 1.5 * IQR
upper_tail <- Q3 + 1.5 * IQR

boxplot(sleep_sample,
        main = "Boxplot of sleep duration",
        ylab = "Sleep duration(hours)")
```

# Bibliography

::: {#refs}
:::

# AI usage

1.  <https://chatgpt.com/share/699e88bf-3d1c-8003-ac39-9186bb8e1d24>
