---
title: "Bootstrap Vignette"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Example Vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
  bibliography: bibliography.bib
nocite: "@*"
---

```{r include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE
)
knitr::opts_chunk$set(fig.width=7, fig.height=6)
```

```{r}
source("bootstrap.R")
```

This vignette demonstrates three functions for sleep duration statistics with the bootstrap resampling method:

-   compute_sleep_duration(): Compute sleep duration as the difference between wake time and bedtime.

-   bootstrap_mean(): Generate bootstrap estimates of the sample mean using resampling with replacement.

-  bootstrap_ci() Compute a percentile bootstrap confidence interval.

# Compute sleep duration

In this example, bedtimes and wake times are recorded on a 24-hour clock.
If wake time is smaller than bedtime, the function assumes sleep crosses midnight and adjusts accordingly.

```{r}
set.seed(1)
n <- 10^4
uptime  <- c( 7.0, 6.0, 9.0, 7.5, 6.0)
bedtime <- c(22.0, 22.5, 1.0, 0.0, 0.5)
sleep_hours <- compute_sleep_duration(bedtime, uptime)
sleep_hours
summary(sleep_hours)
```


# Bootstrap mean of sleep duration

we apply the bootstrap to estimate the mean sleep duration of the sampling distribution.

```{r}
set.seed(1)
B <- 10^4
boot_means_sleep <- bootstrap_mean(sleep_hours, B = B)
head(boot_means_sleep)
length(boot_means_sleep)
```

## Histtogram 

Visualizing the bootstrap of mean sleep duration

```{r}
hist(boot_means_sleep,
     main = "Bootstrap distribution of mean sleep duration",
     xlab = "Bootstrap means(hours)")
```


# Percentile bootstrap confidence interval

Compute the lower and upper bound of the 95% percentile bootstrap confidence interval. 

```{r}
cl95_sleep <- bootstrap_ci(boot_means_sleep, conf_level = 0.95)
cl95_sleep
```

# Bibliography



# AI usage
1. https://chatgpt.com/share/699e88bf-3d1c-8003-ac39-9186bb8e1d24








